
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

####*Data Exploration*

Load the data into R dataframe.


```{r message=FALSE, warning=FALSE, paged.print=FALSE}

library(plyr)
library(dplyr)
library(pastecs)
library(corrplot)
library(caTools)
library(randomForest)
library(caret)
library(party)
library(gridExtra)

setwd("/Volumes/Meghna/Marketing Analytics/Analyses/Customer Churn")
dfCustomer <- read.csv(file="Telco-Customer-Churn.csv", header = TRUE, sep = ",")
print(str(dfCustomer))
```

Check for missing values and drop those rows

```{r warning=FALSE}
sapply(dfCustomer, function(x) sum(is.na(x)))
dfCustomer <- dfCustomer[complete.cases(dfCustomer), ]
```

Replace non-categorical values like "No phone service" with "No" for columns

```{r echo=FALSE, warning=FALSE}
dfCustomer$MultipleLines <- as.factor(mapvalues(dfCustomer$MultipleLines, from=c("No phone service"), to=c("No")))

for(i in 10:15) {
  dfCustomer[,i] <- as.factor(mapvalues(dfCustomer[,i], from =c("No internet service"),to=c("No")))
}
```

Change seniorcitizen from numerical to categorical

```{r}
dfCustomer$SeniorCitizen <- as.factor(mapvalues(dfCustomer$SeniorCitizen,from=c("0","1"),to=c("No", "Yes")))
```

Group "tenure" to change it from continuous to categorical 

```{r}
group_tenure <- function(tenure){
  if (tenure >= 0 & tenure <= 12){
    return('0-12 Month')
  }else if(tenure > 12 & tenure <= 24){
    return('12-24 Month')
  }else if (tenure > 24 & tenure <= 48){
    return('24-48 Month')
  }else if (tenure > 48 & tenure <=60){
    return('48-60 Month')
  }else if (tenure > 60){
    return('> 60 Month')
  }
}
dfCustomer$tenure_group <- sapply(dfCustomer$tenure,group_tenure)
dfCustomer$tenure_group <- as.factor(dfCustomer$tenure_group)

```

Remove the columns we do not need for the analysis.

```{r}
dfCustomer$customerID <- NULL
dfCustomer$tenure <- NULL
print(str(dfCustomer))
```

Convert character data into factors

```{r}
dfCustomer$Churn <- as.factor(dfCustomer$Churn)
dfCustomer$gender <- as.factor(dfCustomer$gender)
dfCustomer$PaperlessBilling <- as.factor(dfCustomer$PaperlessBilling)
dfCustomer$Contract <- as.factor(dfCustomer$Contract)
dfCustomer$Partner <- as.factor(dfCustomer$Partner)
dfCustomer$Dependents <- as.factor(dfCustomer$Dependents)
dfCustomer$PhoneService <- as.factor(dfCustomer$PhoneService)
print(str(dfCustomer))
```

*__Summary Statistics__*

```{r}
numeric_var <- sapply(dfCustomer, is.numeric)
options(scipen=100)
options(digits = 2)
stat.desc(dfCustomer[,numeric_var])
```

Bar plots for categorical variables
```{r}

p1 <- ggplot(dfCustomer, aes(x=gender)) + ggtitle("Gender") + xlab("Gender") +
  geom_bar(aes(y = 100*(..count..)/sum(..count..)), width = 0.3) + ylab("Percentage") + theme_minimal()

p2 <- ggplot(dfCustomer, aes(x=SeniorCitizen)) + ggtitle("Senior Citizen") + xlab("Senior Citizen") + 
  geom_bar(aes(y = 100*(..count..)/sum(..count..)), width = 0.3) + ylab("Percentage") + theme_minimal()

p3 <- ggplot(dfCustomer, aes(x=Partner)) + ggtitle("Partner") + xlab("Partner") + 
  geom_bar(aes(y = 100*(..count..)/sum(..count..)), width = 0.3) + ylab("Percentage")  + theme_minimal()

p4 <- ggplot(dfCustomer, aes(x=Dependents)) + ggtitle("Dependents") + xlab("Dependents") +
  geom_bar(aes(y = 100*(..count..)/sum(..count..)), width = 0.3) + ylab("Percentage") + theme_minimal()
grid.arrange(p1, p2, p3, p4, ncol=2)

p5 <- ggplot(dfCustomer, aes(x=PhoneService)) + ggtitle("Phone Service") + xlab("Phone Service") +
  geom_bar(aes(y = 100*(..count..)/sum(..count..)), width = 0.3) + ylab("Percentage") + theme_minimal()

p6 <- ggplot(dfCustomer, aes(x=MultipleLines)) + ggtitle("Multiple Lines") + xlab("Multiple Lines") + 
  geom_bar(aes(y = 100*(..count..)/sum(..count..)), width = 0.3) + ylab("Percentage") + theme_minimal()

p7 <- ggplot(dfCustomer, aes(x=InternetService)) + ggtitle("Internet Service") + xlab("Internet Service") + 
  geom_bar(aes(y = 100*(..count..)/sum(..count..)), width = 0.3) + ylab("Percentage")  + theme_minimal()

p8 <- ggplot(dfCustomer, aes(x=OnlineSecurity)) + ggtitle("Online Security") + xlab("Online Security") +
  geom_bar(aes(y = 100*(..count..)/sum(..count..)), width = 0.3) + ylab("Percentage") + theme_minimal()
grid.arrange(p5, p6, p7, p8, ncol=2)

p9 <- ggplot(dfCustomer, aes(x=OnlineBackup)) + ggtitle("Online Backup") + xlab("Online Backup") +
  geom_bar(aes(y = 100*(..count..)/sum(..count..)), width = 0.3) + ylab("Percentage")+ theme_minimal()

p10 <- ggplot(dfCustomer, aes(x=DeviceProtection)) + ggtitle("Device Protection") + xlab("Device Protection") + 
  geom_bar(aes(y = 100*(..count..)/sum(..count..)), width = 0.3) + ylab("Percentage")  + theme_minimal()

p11 <- ggplot(dfCustomer, aes(x=TechSupport)) + ggtitle("Tech Support") + xlab("Tech Support") + 
  geom_bar(aes(y = 100*(..count..)/sum(..count..)), width = 0.3) + ylab("Percentage")+ theme_minimal()

p12 <- ggplot(dfCustomer, aes(x=StreamingTV)) + ggtitle("Streaming TV") + xlab("Streaming TV") +
  geom_bar(aes(y = 100*(..count..)/sum(..count..)), width = 0.3) + ylab("Percentage")  + theme_minimal()
grid.arrange(p9, p10, p11, p12, ncol=2)

p13 <- ggplot(dfCustomer, aes(x=StreamingMovies)) + ggtitle("Streaming Movies") + xlab("Streaming Movies") +
  geom_bar(aes(y = 100*(..count..)/sum(..count..)), width = 0.3) + ylab("Percentage") + theme_minimal()

p14 <- ggplot(dfCustomer, aes(x=Contract)) + ggtitle("Contract") + xlab("Contract") + 
  geom_bar(aes(y = 100*(..count..)/sum(..count..)), width = 0.3) + ylab("Percentage")  + theme_minimal()

p15 <- ggplot(dfCustomer, aes(x=PaperlessBilling)) + ggtitle("Paperless Billing") + xlab("Paperless Billing") + 
  geom_bar(aes(y = 100*(..count..)/sum(..count..)), width = 0.3) + ylab("Percentage") + theme_minimal()

p16 <- ggplot(dfCustomer, aes(x=PaymentMethod)) + ggtitle("Payment Method") + xlab("Payment Method") +
  geom_bar(aes(y = 100*(..count..)/sum(..count..)), width = 0.5) + ylab("Percentage") + theme_minimal()

p17 <- ggplot(dfCustomer, aes(x=tenure_group)) + ggtitle("Tenure Group") + xlab("Tenure Group") +
  geom_bar(aes(y = 100*(..count..)/sum(..count..)), width = 0.5) + ylab("Percentage") + theme_minimal()
grid.arrange(p13, p14, p15, p16, p17, ncol=2)
```


####*Feature Selection*

Numeric Variables using correlation matrix
```{r}
corr.matrix <- cor(dfCustomer[,numeric_var])
corrplot(corr.matrix, main="\n\nCorrelation Plot for Numerical Variables", method="number")

dfCustomer$TotalCharges <- NULL #reomve totalcharges because it is correlated to monthlycharges
```

Categorical variables using Pearson's chi-squared test

```{r}
#chisq.test(dfCustomer$Partner, dfCustomer$Churn, correct=FALSE)
```

*__Data Partitioning__*

Sample the input data with 70% for training and 30% for testing

```{r}
set.seed(2017) #Set the seed, it will output same output when ever the model is executed.
sample <- sample.split(dfCustomer$Churn,SplitRatio=0.70)
trainData <- subset(dfCustomer,sample==TRUE)
testData <- subset(dfCustomer,sample==FALSE)
dim(trainData); dim(testData) #Confirm the splitting is correct.
```


####*Model Building*

*__Logistic Regression__*

```{r}
telecomModel <- glm(Churn ~ .,family=binomial(link="logit"),data=trainData)
print(summary(telecomModel))
```

Feature analysis: The top three most-relevant features include Contract, Paperless Billing and tenure group, all of which are categorical variables.

```{r}
anova(telecomModel, test="Chisq")
```

Test the model with test dataset

```{r}
test_prediction <- predict(telecomModel,newdata=testData,type='response')
```

If the prediction probability is greater than 0.5 then those customers are classified as churned customer less than 0.5 are classified as not churning customer

```{r}
fitted_results <- ifelse(test_prediction > 0.5,1,0)
testData$Churn <- as.character(testData$Churn)
testData$Churn[testData$Churn=="No"] <- "0"
testData$Churn[testData$Churn=="Yes"] <- "1"
```

Calculate the misclassfication rate

```{r}
misClasificationError <- mean(fitted_results!=testData$Churn)
```

Calculate the accuracy rate

```{r}
accuracyRate <- 1-misClasificationError
print(paste('Logistic Regression Accuracy',accuracyRate))
```

Confusion matrix

```{r}
print("Confusion Matrix for Logistic Regression"); table(testData$Churn, test_prediction > 0.5)
```

Odds Ratio

```{r}
exp(cbind(OR=coef(telecomModel), confint(telecomModel)))
```

*__Decision Tree__*

```{r fig.height=6, fig.width=10}
tree <- ctree(Churn~Contract+tenure_group+PaperlessBilling, trainData)
plot(tree, type ="simple")
```

Test the model with test dataset

```{r}
pred_tree <- predict(tree, testData)
```

Check for Accuracy

```{r}
p1 <- predict(tree, trainData)
tab1 <- table(Predicted = p1, Actual = trainData$Churn)
tab2 <- table(Predicted = pred_tree, Actual = testData$Churn)
print(paste('Decision Tree Accuracy',sum(diag(tab2))/sum(tab2)))
```

Confusion matrix

```{r}
print("Confusion Matrix for Decision Tree"); table(Predicted = pred_tree, Actual = testData$Churn)
```








